---
title: "R Project - Identifying individuals most likely to click an ad"
author: "Geoffrey Chege"
date: '2022-06-04'
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Introduction

## 1.1 Defining the question
- Determine which individuals are most likely to click on an ad using supervised learning prediction models.

## 1.2 The Context
- A Kenyan entrepreneur has created an online cryptography course and would want to advertise it on her blog.
- She currently targets audiences originating from various countries.
- In the past, she ran ads to advertise a related course on the same blog and collected data in the process.
- She would now like to employ my services as a Data Science Consultant to help her identify which individuals are most likely to click on her ads.

## 1.3 Metric for success
- Accuracy score of 85% and above.

## 1.4 Experimental Design Taken
- Installing packages and loading libraries needed
- Loading the data
- Data Cleaning
- Exploratory Data Analysis:
  - Univariate Analysis
  - Bivariate Analysis
- Modelling
- Predictions and evaluation of the model
- Conclusion

## 1.5 Appropriateness of the available data
- The columns in the dataset include:
  - Daily_Time_Spent_on_Site
  - Age
  - Area_Income
  - Daily_Internet_Usage
  - Ad_Topic_Line
  - City
  - Male
  - Country
  - Timestamp
  - Clicked_on_Ad

# 2. Installing and loading Necessary Packages
```{r, eval=FALSE, include=FALSE}
#install.packages("iterators")  
#install.packages("caret")
#install.packages("caretEnsemble")
#install.packages("ggplot2")
#install.packages("e1071")
#install.packages("randomForest")
#install.packages("ggcorrplot")
#install.packages('ranger')
#install.packages('caTools')
#install.packages('rpart.plot')
#install.packages('psych')
```

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(data.table)
library(dplyr)
library(ggplot2)
library(data.table)
library(mclust)
library(caret)
library(corrplot)
library(clustvarsel)
library(wskm)
library(cluster)
library(reshape2)
library(DataExplorer)
library(lattice)
library(Hmisc)
library(caret)
library(psych)
library(ggcorrplot)
library(tidyr)
library(RColorBrewer)
library("rpart.plot")
```

# 3. Loading the Data


```{r ad}
ad <- read.csv("C:/Users/user/Downloads/advertising.csv") #Loading the dataset
head(ad) #previewing the first 5 elements of the data
```


# 4. Data Cleaning

## 4.1 Checking the attribute types

```{r class, echo=FALSE}
sapply(ad,class) #Checking the attribute types of the columns
```

- The attribute types in the data are: numeric, integer and character.

## 4.2 converting time variable from character to date and time (POSIXct) format
```{r datetime}
ad$Timestamp <- as.POSIXct(ad$Timestamp, "%Y-%m-%d %H:%M:%S",tz = "GMT")
```

## 4.3 Checking for duplicates

```{r duplicates}
duplicates <- ad[duplicated(ad),] #storing duplicates in a table called "duplicates"
duplicates #previewing the table
```

- The duplicates table is empty. This means that there are no duplicates in the dataset.

## 4.4 checking for null values

```{r nulls}
colSums(is.na(ad)) #Checking the total number of null values in each column
```

- There are no null values in the dataset

## 4.5 checking column names

```{r names}
names(ad) #Displaying column names
```

- The data set has the above column names. Columns with more than one word have periods "." separating the words. I will replace the periods "." with underscores "_"

```{r underscore}
names(ad) <- gsub("[.]", "_", names(ad)) #Replacing "." with "_"
```

- The above code replaces the periods "." with underscores "_".

```{r name check}
names(ad) #Displaying column names
```

## 4.6 Outliers

- I will use boxplots to check for outliers.

### Boxplot for "Area_Income"

```{r boxplot income, echo=FALSE}
boxplot(ad$Area_Income,
main = "Area Income Boxplot",  # Titling the boxplot
xlab = "Area Income",          # labelling the axis
col = "blue",                  # setting the colour
border = "black",              # setting the colour of the borders
horizontal = TRUE,
notch = TRUE
)
```

- There are few outliers in the "Area_Income" column. I will not remove them because they will be relevant in the analysis.

### Boxplot for "Age"

```{r boxplot age, echo=FALSE}
boxplot(ad$Age,
main = "Age Boxplot",        # Titling the boxplot
xlab = "Age",                # labelling the axis
col = "orange",              # setting the colour
border = "black",            # setting the colour of the borders
horizontal = TRUE,
notch = TRUE
)
```

- There are no outliers in the "Age" column.

### Boxplot for "Daily_Time_Spent_on_Site"

```{r boxplot time on site, echo=FALSE}
boxplot(ad$Daily_Time_Spent_on_Site,
main = "Time spent on site Boxplot",    # Titling the boxplot
xlab = "Time spent on site",            # labelling the axis
col = "brown",                         # setting the colour
border = "black",                      # setting the colour of the borders
horizontal = TRUE,
notch = TRUE
)
```

- There are no outliers in the "Time_Spent_on_Site" column.

### Boxplot for "Daily_Internet_Usage"

```{r internet usage, echo=FALSE}
boxplot(ad$Daily_Internet_Usage,
main = "Daily Internet usage Boxplot",      # Titling the boxplot
xlab = "Daily internet usage",              # labelling the axis
col = "red",                                # setting the colour
border = "black",                           # setting the colour of the borders
horizontal = TRUE,
notch = TRUE
)
```

- There are no outliers in the "Daily_Internet_Usage" column.

# 5. Exploratory Data Analysis

## 5.1 Univariate Analysis

- Summary statistics of the dataset

```{r summary}
summary(ad)
```

- Using "describe()" function to get range, skewness, kurtosis and standard deviation. The "summary()" function does not give us this information.

```{r inform}
describe(ad)
```

From the "summary()" and "describe()" functions, the following measures of central tendency can be gathered:

### Daily_Time_Spent_on_Site:

- mean: 65
- median: 68.22
- maximum: 91.43
- minimum: 32.60
- range: 58.83
- skew: -0.37
- kurtosis: -1.10

### Age:

- mean: 36.01
- median: 35
- maximum: 61
- minimum: 19
- range: 42
- skew: 0.48
- kurtosis: -0.41

### Area Income:

- mean: 55,000
- median: 57,012
- maximum: 79,484.8
- minimum: 13,996.5
- range: 65,488.30
- skew: -0.65
- kurtosis: -0.11

### Daily_Internet_Usage:

- mean: 180
- median: 183.1
- maximum: 269.96
- minimum: 104.78
- range: 165.18
- skew: -0.03
- kurtosis: -1.28

### Mode

- A function to determine the mode:

```{r mode}
mode <- function(v){
  uniq <- unique(v)
  uniq[which.max(tabulate(match(v,uniq)))]
}
```

The most recurrent Ad Topic Line:
```{r modal topic, echo=FALSE}
topic_mode <- mode(ad$Ad_Topic_Line)
topic_mode
```

The most recurrent City:
```{r modal city, echo=FALSE}
city_mode <- mode(ad$City)
city_mode
```

The most recurrent Country:
```{r modal country, echo=FALSE}
country_mode <- mode(ad$Country)
country_mode
```

- Checking the modal age using a barplot:

```{r age plot, echo=FALSE}
counts <- table(ad$Age)
barplot(counts,
  main="A bar chart showing Age",
  xlab="Age",
  ylab = "Frequency",
  col=c("darkblue"),
)
```

- From the plot, the modal age is 31.


- Checking the distribution in terms of gender where 1 is Male and 0 is Female:

```{r gender distribution, echo=FALSE}
gender <- (ad$Male)
gender.frequency <- table(gender)
gender.frequency
# plotting to visualize the distribution
barplot(gender.frequency,
  main="A bar chart showing Gender",
  xlab="Gender",
  ylab = "Frequency",
  col=c("darkblue","red"),
  )
```

From this, there are More women than men, making female the modal gender.

## 5.2 Bivariate Analysis

### Scatterplots

```{r}
# scatterplot
plot((ad$Daily_Time_Spent_on_Site), (ad$Age),
     main = "A scatterplot of Time Spent on site against age",
     xlab = 'Time spent',
     ylab = 'Age')
```

```{r}
# scatterplot of Time on site vs income
plot((ad$Daily_Time_Spent_on_Site), (ad$Area_Income),
     main = "A scatterplot of Time Spent on site against income",
     xlab = 'Time Spent on Site',
     ylab = 'Income')
```

```{r}
# scatterplot of Time on site vs Internet usage
plot((ad$Daily_Time_Spent_on_Site), (ad$Daily_Internet_Usage),
     main = "A scatterplot of Time Spent on site against Daily Internet Usage",
     xlab = 'Time Spent on Site',
     ylab = 'Daily Internet Usage')
```

### Heatmap

```{r heatmap}
# Heat map
# Checking the relationship between the variables
# Using Numeric variables only
numeric_tbl <- ad %>%
  select_if(is.numeric) %>%
  select(Daily_Time_Spent_on_Site, Age, Area_Income,Daily_Internet_Usage)
# Calculate the correlations
corr <- cor(numeric_tbl, use = "complete.obs")
ggcorrplot(round(corr, 2),
           type = "full", lab = T)
```

### Analysis of those who clicked on ads:

```{r}
# Analysis of people who click on the ads
ad_click <- ad[which(ad$Clicked_on_Ad == 1),] # Creating a new dataset that only has those who clicked on an ad
```

- Most popular age group of people clicking on ads:

```{r}
# Most popular age group of people clicking on ads
hist((ad_click$Age),
     main = "Histogram of Age of those who click ads",
     xlab = 'Age',
     ylab = 'Frequency',
     col = "blue")
```

- 40 - 45 year olds click on the most ads.

### Plotting to visualize the gender distribution:

```{r}
gender2 <- (ad_click$Male)
gender2.frequency <- table(gender2)
gender2.frequency
# plotting to visualize the gender distribution
barplot(gender2.frequency,
  main="A bar chart showing Gender of those who clicked",
  xlab="Gender(0 = Female, 1 = Male)",
  ylab = "Frequency",
  col=c("darkblue","red"),
  )
```

- Females clicked more ads than males.

### Scatterplots of those who clicked:

```{r}
# scatterplot
plot((ad_click$Daily_Time_Spent_on_Site), (ad_click$Age),
     main = "A scatterplot of Time Spent on site and clicked ad against age",
     xlab = 'Time spent',
     ylab = 'Age')
```

```{r}
# scatterplot of Time on site vs income
plot((ad_click$Daily_Time_Spent_on_Site), (ad_click$Area_Income),
     main = "A scatterplot of Time Spent on site and ad clicked against income",
     xlab = 'Time Spent on Site',
     ylab = 'Income')
```

```{r}
# scatterplot of Time on site vs Internet usage
plot((ad_click$Daily_Time_Spent_on_Site), (ad_click$Daily_Internet_Usage),
     main = "A scatterplot of Time Spent on site and ad clicked against Daily Internet Usage",
     xlab = 'Time Spent on Site',
     ylab = 'Daily Internet Usage')
```


```{r heatmap clicked}
# Heat map
# Checking the relationship between the variables

# Using Numeric variables only
numeric_tbl <- ad_click %>%
  select_if(is.numeric) %>%
  select(Daily_Time_Spent_on_Site, Age, Area_Income,Daily_Internet_Usage)

# Calculate the correlations
corr <- cor(numeric_tbl, use = "complete.obs")
ggcorrplot(round(corr, 2),
           type = "full", lab = T)
```

- There is low correlation between the numerical variables.

- The country with the most ad clicks:

```{r}
mode(ad_click$Country)
```

- The income that clicks most:

```{r}
mode(ad_click$Area_Income)
```

- Ad title that garners most clicks:

```{r modal topic clicked, echo=FALSE}
topic_mode2 <- mode(ad_click$Ad_Topic_Line)
topic_mode2
```


- All the data profiling statistics will be organized into the report below

```{r}
create_report(ad)
```

- A link to the report: "https://github.com/Geoffrey-Chege/Supervised-and-Unsupervised-Learning/blob/main/Ad%20Clicks/report.html"


# 6. Modelling

```{r heatmap all numeric}
# Heat map
# Checking the relationship between the variables

# Using Numeric variables only
numeric_tbl2 <- ad %>%
  select_if(is.numeric) %>%
  select(Daily_Time_Spent_on_Site, Age, Area_Income,Daily_Internet_Usage, Clicked_on_Ad)

# Calculate the correlations
corr <- cor(numeric_tbl2, use = "complete.obs")
ggcorrplot(round(corr, 2),
           type = "full", lab = T)
```

1. Daily_Time_Spent_on_Site and Clicked_on_Ad variables are strongly inversely related with a correlation of -0.75.
2. Daily_Internet_Usage and Clicked_on_Ad are strongly variable are strongly inversely related with a correlation of - 0.79.
3. Daily_Time_Spent_on_Site and Daily_Internet_Usage variables are positively related with 0.52. correlation.
4. Age and Daily_Internet_Usage variables are positively related with 0.49 correlation.

Clicked_on_Ad is the target variable so I will get correlation without it included.

```{r heatmap all numeric except Clicked_on_Ad}
# Heat map
# Checking the relationship between the variables

# Using Numeric variables only
numeric_tbl3 <- ad %>%
  select_if(is.numeric) %>%
  select(Daily_Time_Spent_on_Site, Age, Area_Income,Daily_Internet_Usage)

# Calculating the correlations
corr <- cor(numeric_tbl3, use = "complete.obs")
ggcorrplot(round(corr, 2),
           type = "full", lab = T)
```

- There are no highly correlated numeric independent variables, so I will use them all in analysis.

## Normalizing the independent variables to ensure all the data is on the same scale

```{r}
# Normalizing the dataset
normalize <- function(x){
  return ((x-min(x)) / (max(x)-min(x)))
}
ad$Daily_Time_Spent_on_Site <- normalize(ad$Daily_Time_Spent_on_Site)
ad$Age <- normalize(ad$Age)
ad$Area_Income <- normalize(ad$Area_Income)
ad$Male <- normalize(ad$Male)

#previewing normalized dataset
head(ad)
```

- The dataset is on the same scale.

## Splitting Data into Training and Testing Sets

```{r}
# splitting the data into training and testing sets
# I will split it 70:30
intrain <- createDataPartition(y = ad$Clicked_on_Ad, p = 0.7, list = FALSE)
training <- ad[intrain,]
testing <- ad[-intrain,]
```



```{r}
# checking the dimensions of our training and testing sets
dim(training)
dim(testing)
```

- 700 of data will be used for training while 300 will be for testing.

```{r}
# checking the dimensions of our split
prop.table(table(ad$Clicked_on_Ad)) * 100
prop.table(table(training$Clicked_on_Ad)) * 100
prop.table(table(testing$Clicked_on_Ad)) * 100
```

- The target data is equal in the data,training set and test set.


## Decision Tree Classifier

```{r}
# Specifying target and predictor variables
m <- rpart(Clicked_on_Ad ~ . ,
         data = numeric_tbl2,
         method = "class")
```

```{r}
# Plotting model
rpart.plot(m)
```

```{r}
# Making predictions
p <- predict(m, numeric_tbl2, type ="class")

# Printing the confusion matrix
table(p, numeric_tbl2$Clicked_on_Ad)
```

- The model correctly classified 485 did not clicks as ‘0’ and 472 clicks as ‘1’ . However, it also incorrectly classified 28 did not clicks as ‘1’(clicked) and 15 clicks as ‘0’(did not click).

```{r}
# Printing the Accuracy
(mean(numeric_tbl2$Clicked_on_Ad == p))*100
```

  - The model has an accuracy of 95.7%
  - This is a good model for making predictions

# 7. Conclusions

- Decision Tree gives an accuracy of 95.7%
- The females have the majority site visits but they don't click on the ad.
- The minimum age of the participant was 19 years old while the oldest was 60 years old.
- The minimum daily time spent on the site was 32 minutes while the maximum time spent was 91 minutes.
- The youth have most site visits as compared to the teenagers and older people.

# 8. Recommendations

- Appropriate content targeting different age groups should be uploaded when it comes to the ads. This will lead to an increase in the number of clicks on ads.
- There should be more locally targeted ads, seeing as the key word 'local' prompted more clicks.